### 2019-04-18 figures/final/datasets.tex
Update the whole paper, particularly the results

\begin{tabular}{|p{0.18\linewidth}|p{0.14\linewidth}|p{0.1\linewidth}|p{0.15\linewidth}|p{0.23\linewidth}|p{0.16\linewidth}|}
  \hline
  Dataset         & Dimensions   & Number of Batches & Batch Label      & True Class Label       & Data Type        \\
  \hline
  Bladder Batch   & $57\times22,283$  & 5                 & Batch            & Cancer status          & Microarray       \\
  \hline
  GSE37199        & $93\times20,024$  & 2                 & Plate            & Cancer stage           & Microarray       \\
  \hline
  MNIST           & $10,000\times784$ & 2                 & Artificial batch & Digit                  & Grayscale images \\
  \hline
  TCGA Pan-Cancer & $9,366\times325$  & 25                & Cancer Type      & TP53 mutation presence & RNA-Seq \\
  \hline
\end{tabular}