### 2019-07-11 figures/code/loss_chart.Rmd
Add all the code that was removed from Confounded

---
title: "Loss Charts"
author: "Jonathan Dayton"
date: "12/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load("tidyverse", "gridExtra", "png", "stringr", "magick", "colorspace", "pracma")
```

# Load data

```{r}
training_log_files <- dir("../../data/metrics/training/", pattern = ".*\\.csv", full.names = TRUE)
log_dfs <- do.call("rbind", lapply(training_log_files, read_csv))
```

# Look at the data

## Number of autoencoder layers doesn't seem to matter after 1

```{r}
final_iter <- log_dfs %>% group_by(start_time) %>% summarize(iteration = max(iteration)) %>% left_join(log_dfs)
final_iter %>% filter(str_detect(output_path, "ae_layers_")) %>%
  ggplot(aes(x = autoencoder_layers, y = dual_loss)) +
  geom_line() +
  labs(x = "Number of Autoencoder Layers", y = "Dual Loss Value")
```

## More layers in the discriminator generally means lower final loss

```{r}
final_iter %>% filter(str_detect(output_path, "/layers_")) %>%
  filter(discriminator_layers > 0) %>%
  ggplot(aes(x = discriminator_layers, y = dual_loss)) +
  geom_line() +
  labs(x = "Number of Discriminator Layers", y = "Dual Loss Value")
```

## The optimal code size seems to be between 1000 and 10,000

```{r}
final_iter %>% filter(str_detect(output_path, "/code_size")) %>%
  filter(discriminator_layers > 0) %>%
  ggplot(aes(x = code_size, y = dual_loss)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Nodes in Code Layer", y = "Dual Loss Value") +
  scale_x_log10()
```

```{r}
final_iter %>% filter(str_detect(output_path, "/scaling")) %>%
  ggplot(aes(x = scaling_method, y = dual_loss)) +
  geom_bar(stat = "identity") +
  labs(x = "Scaling Method", y = "Dual Loss Value")
```

[1] "../../data/metrics/training//ae_layers.csv"   "../../data/metrics/training//code_size.csv"  
[3] "../../data/metrics/training//features.csv"    "../../data/metrics/training//layers.csv"     
[5] "../../data/metrics/training//loss_weight.csv" "../../data/metrics/training//mnist.csv"      
[7] "../../data/metrics/training//scaling.csv"

```{r}

run <- log_dfs %>% filter(discriminator_layers == 9)

run %>%
  # filter(iteration > 4000) %>%
  ggplot(aes(x = iteration, y = dual_loss)) +
  geom_line() +
  geom_smooth() +
  labs(title = "Dual Loss", y = "Loss", x = "Iteration")

run %>%
  ggplot(aes(x = iteration, y = ae_loss)) +
  geom_line() +
  geom_smooth() +
  labs(title = "Autoencoder Loss", y = "Loss", x = "Iteration")

run %>%
  ggplot(aes(x = iteration, y = disc_loss)) +
  geom_line() +
  geom_smooth() +
  labs(title = "Discriminator Loss", y = "Loss", x = "Iteration")
```