### 2019-07-16 code/figures/code/mnist.Rmd

---
title: "MNIST"
author: "Jonathan Dayton"
date: "7/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load("tidyverse", "gridExtra", "png", "stringr", "magick", "colorspace", "pracma")
```

## MNIST Grid

```{r}
mnist_dir <- "~/Documents/research/confounded/data/mnist/"

dfs <- list(
  Unadjusted = read_csv(paste(c(mnist_dir, "unadjusted.csv"), collapse="")) %>% select(-Batch, -Sample, -Digit),
  `Artificial noise` = read_csv(paste(c(mnist_dir, "noisy.csv"), collapse="")) %>% select(-Batch, -Sample, -Digit),
  Scale = read_csv(paste(c(mnist_dir, "noisy_scale.csv"), collapse="")) %>% select(-Batch, -Sample, -Digit),
  ComBat = read_csv(paste(c(mnist_dir, "noisy_combat.csv"), collapse="")) %>% select(-Batch, -Sample, -Digit),
  Confounded = read_csv(paste(c(mnist_dir, "noisy_confounded.csv"), collapse="")) %>% select(-Batch, -Sample, -Digit)
)
```

```{r}
unadj <- dfs$Unadjusted
set.seed(8)
pic_rows <- sort(sample.int(nrow(unadj), 5))
print(pic_rows)
pic_rows <- c(1, 2223, 4445, 1112, 3334, 5556)

image_from_vector <- function(x) {
  return(image_read(aperm(array(as.numeric(x), c(28, 28, 1)), c(2, 1, 3))))
}

big_img <- NULL
imgs <- list()
for (name in names(dfs)) {
  img <- NULL
  df <- dfs[[name]]
  for (i in pic_rows) {
    new_img <- image_from_vector(df[i,])
    if (is.null(img)) {
      img <- new_img
    } else {
      img <- image_append(c(img, new_img))
    }
  }
  imgs[[i]] <- img
  if (is.null(big_img)) {
    big_img <- img
  } else {
    big_img <- image_append(c(big_img, img), stack = TRUE)
  }
}

label_width <- 65
text_background <- image_scale(image_read(array(rep(1, 28*label_width), c(28, label_width, 1))), "x100")
labels <- NULL
for (name in names(dfs)) {
  new_label <- image_annotate(text_background, name, gravity = "center", size = 32)
  if (is.null(labels)) {
    labels <- new_label
  } else {
    labels <- image_append(c(labels, new_label), stack = TRUE)
  }
}

big_img <- image_scale(big_img, "500")
labelled <- image_append(c(labels, big_img))

image_write(labelled, "../../../../research/thesis/figures/final/mnist.png")
```